---
title: "GISD Final Project"
author: "Anna Church"
date: "11/12/2019"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load('GISD_project_data.RData')
```

# Infant Mortality Rates & Household Income
### Publicly available from CDC WONDER linked birth/infant death files from 2011-2015 & Area Health Resource File county-level variables. I linked these files by county FIPS codes. Links to these raw data are available in my LaTex file, and I have provided the cleaned data in my GitHub repository. 

# Descriptive Statistics

```{r}
# summarize the data
summary(Thesis_Final_Data_Set_2)
summary(Thesis_Final_Data_Set_2$BLACK_IMR)
summary(Thesis_Final_Data_Set_2$WHITE_IMR)
# means by themselves
mean(Thesis_Final_Data_Set_2$BLACK_IMR)
mean(Thesis_Final_Data_Set_2$WHITE_IMR)
# histograms
hist(Thesis_Final_Data_Set_2$BLACK_IMR)
hist(Thesis_Final_Data_Set_2$WHITE_IMR)
```

# Scatterplot of Black and White Infant Mortality Rates (IMR)

```{r}
plot(Thesis_Final_Data_Set_2$BLACK_IMR, Thesis_Final_Data_Set_2$WHITE_IMR, main = 'Scatterplot',
     xlab= 'Black IMR', ylab = 'White IMR',
     col= 'blue', pch= 3, cex= 2)
```

# Linear Regression of Black IMR on Household Income

```{r}
model.1 <- lm(Thesis_Final_Data_Set_2$BLACK_IMR ~ Thesis_Final_Data_Set_2$V_med_hhld_income)
# produces regression results
summary(model.1)
# store the predicted values
model.1.pred <- predict(model.1)
# first data
plot(x=Thesis_Final_Data_Set_2$V_med_hhld_income,y=Thesis_Final_Data_Set_2$BLACK_IMR,xlab="Median Household Income",ylab="Black IMR")
# now predicted values
points(x=Thesis_Final_Data_Set_2$V_med_hhld_income,y=model.1.pred,type="l",col="red")
# the predicted values are also in model.1 ...
points(x=Thesis_Final_Data_Set_2$V_med_hhld_income,y=model.1$fitted.values,col="blue",cex=0.5)
```

# Linear Regression of White IMR on Household Income

```{r}
model.2 <- lm(Thesis_Final_Data_Set_2$WHITE_IMR ~ Thesis_Final_Data_Set_2$V_med_hhld_income)
# produces regression results
summary(model.2)
# store the predicted values
model.2.pred <- predict(model.2)

# plot predictions for the most interesting relationship 
# first data
plot(x=Thesis_Final_Data_Set_2$V_med_hhld_income,y=Thesis_Final_Data_Set_2$WHITE_IMR,xlab="Median Household Income",ylab="White IMR")
# now predicted values
points(x=Thesis_Final_Data_Set_2$V_med_hhld_income,y=model.2.pred,type="l",col="red")
# the predicted values are also in model.1 ...
points(x=Thesis_Final_Data_Set_2$V_med_hhld_income,y=model.2$fitted.values,col="blue",cex=0.5)
```
